NURM: a tool to locate numerical differences in pipelines

### 2019-06-04 paper.tex
Updates

Figure~\ref{fig:overview} shows an overview of our method. For
each tested condition, the pipeline is containerized with Docker, and its
interface is described with Boutiques~\cite{glatard2017boutiques}.
\reprozip~\cite{Chirigati2016} is added to one of the container images for
provenance tracking. 
Our tool starts by capturing temporary and multi write files in Condition 1 
(Fig.~\ref{fig:overview-capturing}). 
To do that, the tool first executes pipeline on Condition 1 to obtain process graph 
(Fig.~\ref{fig:overview-capturing} - (P)) and 
result files. Subsequently, it generates the list of processes that create these intermediary files. 
Using this list, the tool modifies docker image in Condition 1 and exectutes the pipeline 
again through it, which turns to produce results including temporary and overwritten
files (Fig.~\ref{fig:overview-capturing} - (C1)).
In the second step, the processes are labeled from 
the differences found in their outputs in the
two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order, 
the tool builds docker image for Condition 2 by modifying pipeline processes that 
recognised from 
the process graph. Then, the modified docker image is executed in Condition 2 and 
each process is labeled right 
after the running by comparing checksum of the results obtained in Condition 1 
and output files of the process in Condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
We then label processes in 2 categories (Fig.~\ref{fig:overview-labelling} -
(L)) depending whether they create differences (represented in red) or not (green). 
Process labelling is done through hooks which are defined during the docker image modifying procedure, 
and incremental updates in Condition 2, 
until the end of the pipeline execution.

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=1\columnwidth]{images/Slide2}
    \caption{Capturing of temp and multi write files.}
    \label{fig:overview-capturing}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=1\columnwidth]{images/Slide1}
     \caption{Process labelling.}
     \label{fig:overview-labelling}
   \end{subfigure}
   \caption{Method overview}
   \label{fig:overview}
  \end{figure}

@@

We use \reprozip to create process graph (Fig.~\ref{fig:overview-capturing} - (P)), 
which records: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call and stores it in a SQLite database.

Our tool reconstructs a process tree starting from the first process
created by the pipeline and traversing processes through \texttt{clone()}
and \texttt{fork()}. It also creates a process \emph{graph} from this tree
by adding edges corresponding to file dependencies between processes. A
file dependency is defined between processes A and B if a file written by A
is read by B. Figure~\ref{fig:simple_script} shows an example of a process
tree and graph constructed from the example pipeline in
Listing~\ref{listing:sample-script}.
\begin{listing}
\begin{minted}[frame=single,
  framesep=3mm,
  linenos=false,
  xleftmargin=0pt,
  tabsize=4]{bash}
#!/bin/bash
if [ #$ !=1 ]
then
    echo "usage: 0 <inputimage.nii.gz>"
    exit 1
fi

# Parse argument, set file names
input_image=$1
bet_output="$(basename ${input_image} 
                       .nii.gz)_brain.nii.gz"
bet_output_binarized="$(basename ${input_image} 
                       .nii.gz)_brain_bin.nii.gz"

# Run FSL bet, put result ${bet_output}
bet ${input_image} ${bet_output} > bet_temp.out
echo "Voxels / Volume in brain mask:"
# Run FSL stats on ${bet_output}
fslstats ${bet_output} -V
# Run FSL maths on ${bet_output}
fslmaths ${bet_output} ${bet_output_binarized}
echo "Voxels / Volume in binarized brain mask:"

\rm bet_temp.out
\end{minted}
  \caption{Example pipeline}
  \label{listing:sample-script}
\end{listing}

\begin{figure}
\centering
  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in yellow and result files are in green.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{bash} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. 
}
  \label{fig:simple_script}
\end{figure}

@@

We label in two categories the processes of the graph of each subject:
\begin{enumerate}
\item Processes that read files that do not have differences and write files that do not have differences 
are labeled \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have differences but write files that have differences 
\emph{create} differences (Figure~\ref{fig:processes}.b).
\end{enumerate}

In addition, there are two more situations in labelling process as:
\begin{enumerate}
\item Processes that read files 
that have differences and write files that do not have differences \emph{remove} 
differences (Figure~\ref{fig:processes}.c).
\item Processes that read files that have differences and write files that 
also have differences are called
\emph{undetermined} (Figure~\ref{fig:processes}.d).
\end{enumerate}

\begin{figure}%\centering
\centering
    \begin{subfigure}{0.2\linewidth}
        \includegraphics[scale=0.3]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/red.png}
    \caption{Creates differences}
    \label{fig:red}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/blue.png}
    \caption{Removes differences}
    \label{fig:blue}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/yellow.png}
    \caption{Undetermined}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of labeled process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B. Solid black edges refer to the 
  relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

Undetermined processes are due to temporary files, and files that were
written by more than a single process. In both situations, hooks must
be added to the pipeline to capture intermediary file states and label
undetermined processes.

@@

The labelling of processes that read or 
write temporary files is undetermined because these files are deleted during 
the execution. To address this issue, we replace in Condition 1 every process P that 
writes temporary files with a modification of P that first calls P and then
backs up all its output files to a read-only directory. This replacement is
done by modifying the PATH environment variable to point to a directory
containing a modified version of the executable called by the process. The
modified executable will run the original executable and then backup its
output files (see Listing~\ref{listing:modify-script}). 
This 
solution does not cover the temporary files that are removed by P 
itself, but this is not a problem since these files do not play any role in 
the subsequent steps of the pipeline.

@@

Files written by multiple processes also lead 
to undetermined labels. For a file F 
written by all processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
(1) check that processes in \textbf{P} do not write concurrently to F, 
(2) we establish an order on \textbf{P} based on the creation timestamp 
of the processes, (3) we replace every process $P_{i}$ in \textbf{P} by 
a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
directory. Thus, multiple versions of F are saved and used in the 
analysis.

@@

\begin{listing}
\begin{minted}[frame=single,
  framesep=3mm,
  linenos=false,
  xleftmargin=0pt,
  tabsize=4]{python}
#!/bin/python

# Execute process in cond2 from backup folder
p_name = sys.argv[0]
p_exec_cond2 = op.join('backup_scripts', p_name)
i = 1
while i < len(sys.argv):
    p_exec_cond2 += " " + sys.argv[i]
    i += 1
subprocess.Popen(p_exec_cond2, shell=True).communicate()

# Load list of process to be checked
with open('path/to/p_list/', 'r') as p_list:
    process_list = json.load(p_list)
for p_exec_cond1, p_files in process_list.items():
    # Comparing process exec in two conditions
    if p_exec_cond1 == p_exec_cond2:
        # Make backup copy
        capture_file(p_files, from_cond1, to_backup)

\end{minted}
  \caption{Modify script.}
  \label{listing:modify-script}
\end{listing}

>>>>

Figure~\ref{fig:overview} shows an overview of our method. For each tested
condition, Condition 1 and Condition 2, the pipeline is containerized with
Docker, and its interface is described with
Boutiques~\cite{glatard2017boutiques}. \reprozip~\cite{Chirigati2016} is
added to Condition 1 for provenance tracking. Our tool starts by
identifying temporary files and ``multi-write" files, i.e., files written
by more than one process (Fig.~\ref{fig:overview-capturing}) \tristan{check if concurrent writes are discussed}. To do that,
the tool first executes the pipeline in Condition 1, to obtain a process
graph (Fig.~\ref{fig:overview-capturing} - (P)) and a list of result files.
The list of processes that create intermediary or multi-write
files is then generated. Using this list, the tool modifies Docker image in Condition 1 and
exectutes the pipeline again through it, which turns to produce results
including temporary and overwritten files
(Fig.~\ref{fig:overview-capturing} - (C1)). In the second step, the
processes are labeled from the differences found in their outputs in the
two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order,
the tool builds docker image for Condition 2 by modifying pipeline
processes that recognised from the process graph. Then, the modified docker
image is executed in Condition 2 and each process is labeled right after
the running by comparing checksum of the results obtained in Condition 1
and output files of the process in Condition 2
(Fig.~\ref{fig:overview-labelling} - (D)). We then label processes in 2
categories (Fig.~\ref{fig:overview-labelling} - (L)) depending whether they
create differences (represented in red) or not (green). Process labelling
is done through hooks which are defined during the docker image modifying
procedure, and incremental updates in Condition 2, until the end of the
pipeline execution.

\tristan{Fig 1-a isn't in the git repo while I had asked it before and it is 
written in the getting started guide anyway. It is annoying to have to ask for these things
multiple times while I would like to focus on the content. Also, ppt is among the worst formats
to put in a git repo as it doesn't play well with diffs.}

\tristan{In Fig 1-a, change ``overwritten'' to ``multi-write''.}

\tristan{We should also discuss the cardinality of ``input subject''. Are we talking
about 1 subject or all subjects?}

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=1\columnwidth]{images/Slide2}
    \caption{Identification of temporary and multi-write files.}
    \label{fig:overview-capturing}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=1\columnwidth]{images/Slide1}
     \caption{Process labelling.}
     \label{fig:overview-labelling}
   \end{subfigure}
   \caption{Method overview}
   \label{fig:overview}
  \end{figure}

@@

We use \reprozip (Fig.~\ref{fig:overview-capturing} - (P))
to capture: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call and stores it in a SQLite database.

Our tool reconstructs a process tree starting from the first process
created by the pipeline and traversing processes through \texttt{clone()}
and \texttt{fork()} dependencies. Our tool also creates a process \emph{graph} from this tree
by adding edges corresponding to file dependencies between processes. A
file dependency is defined between processes A and B if a file written by A
is read by B. Figure~\ref{fig:simple_script} shows an example of a process
tree and graph constructed from the example pipeline in
Listing~\ref{listing:sample-script}.
\begin{listing}
\begin{minted}[frame=single,
  framesep=3mm,
  linenos=false,
  xleftmargin=0pt,
  tabsize=4]{bash}
#!/bin/bash

# This script extracts the brain from an input image,
# measures the volume of the extracted brain,
# and creates a binarized brain mask.

if [ $# != 1 ]
then
    echo "usage: $0 <inputimage.nii.gz>"
    exit 1
fi

# Parse argument, set file names
input_image=$1
bet_output=$(basename "${input_image}" \
                       .nii.gz)_brain.nii.gz
bet_output_binarized=$(basename "${input_image}" \
                       .nii.gz)_brain_bin.nii.gz

# Run FSL bet, put result ${bet_output}
bet "${input_image}" "${bet_output}" > bet_temp.out
echo "Voxels / Volume in brain mask:"
# Run FSL stats on ${bet_output}
fslstats "${bet_output}" -V
# Run FSL maths on ${bet_output}
fslmaths "${bet_output}" "${bet_output_binarized}"
# Remove temporary file
\rm bet_temp.out
\end{minted}
  \caption{Example pipeline}
  \label{listing:sample-script}
\end{listing}

\begin{figure}
\centering
  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in yellow and result files are in green.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. 
}
  \label{fig:simple_script}
\end{figure}

@@

We label in two categories the processes of the graph of each subject. First,
processes that read files that do not have differences and write files that
do not have differences are labeled \emph{transparent}
(Figure~\ref{fig:green}). Second, processes that read files that do not
have differences but write files that have differences \emph{create}
differences (Figure~\ref{fig:red}). In addition, processes that read files
that have differences and write files that also have differences are called
\emph{undetermined} (Figure~\ref{fig:yellow}). Processes are also labeled
undetermined when they read or write temporary files, or produce multi-write
files. To resolve undetermined processes, hooks must be added to the pipeline to 
execute these processes on files without differences, and to capture
intermediary file states.

\begin{figure}%\centering
\centering
    \begin{subfigure}{0.2\linewidth}
        \includegraphics[scale=0.3]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/red.png}
    \caption{Creates differences}
    \label{fig:red}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/yellow.png}
    \caption{Undetermined}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of labeled process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B. Solid black edges refer to the 
  relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

\tristan{You should use Fig. or Figure consistently}

@@

The label of processes that read or 
write temporary files is undetermined because these files are deleted during 
the execution. To address this issue, we replace in Condition 1 every process P that 
writes temporary files with a modification of P that first calls P and then
saves all its output files to a read-only directory. This replacement is
done by modifying the PATH environment variable to point to a directory
containing a modified version of the executable called by the process. The
modified executable will run the original executable and then save its
output files (see Listing~\ref{listing:modify-script}). 
\tristan{Add more comments, in particular to explain what the variables represent.}
This 
solution does not cover the temporary files that are removed by P 
itself, but this is not a problem since these files do not play any role in 
the subsequent steps of the pipeline. 
\begin{listing}
  \begin{minted}[frame=single,
    framesep=3mm,
    linenos=false,
    xleftmargin=0pt,
    tabsize=4]{python}
  #!/usr/bin/env python
  import os.path as op

  # Execute process in cond2 from backup folder
  p_name = sys.argv[0]
  p_exec_cond2 = op.join('backup_scripts', p_name)
  p_exec_cond2 = p_exec_cond2 + " ".join(sys.argv[1:])
  subprocess.Popen(p_exec_cond2, shell=True).communicate()

  # Load list of process to be checked
  with open('path/to/p_list/', 'r') as p_list:
      process_list = json.load(p_list)

  if p_exec_cond1 in process_list.keys():
      # Make backup copy
      capture_file(p_files, from_cond1, to_backup)

  \end{minted}
    \caption{Modified version of the script in Listing~\ref{listing:sample-script} to save temporary files.}
    \label{listing:modify-script}
  \end{listing}

@@

Files written by multiple processes also lead 
to undetermined labels. For a file F 
written by all processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
(1) check that processes in \textbf{P} do not write concurrently to F, 
(2) we establish an order on \textbf{P} based on the creation timestamp 
of the processes, (3) we replace every process $P_{i}$ in \textbf{P} by 
a wrapper that first calls $P_{i}$ and then saves F to a read-only 
directory. Thus, multiple versions of F are saved and used in the 
analysis.

@@
